\tool{}: Specification- and Syntax-based Automated Testing Linguistic Capabilities of NLP Models

### 2023-01-13 paper/fse23/intro.tex
intro

To address this limitation, methodologies to assess an NLP model on a set of
  linguistic capabilities have been introduced~\cite{marcoACL2020checklist,rottger2020hatecheck}. 
A \lc
explains a functionality of the input and output behavior for the
NLP model under test. Typically, it describes certain type of inputs
and outputs observed in real world for the target NLP task.
% , ranging
% from simple to complicated behaviors.
Testing the linguistic capabilities allows the model developers to better understand the capabilities and potential issues of the NLP
models. 
For example,  \Cklst~\cite{marcoACL2020checklist}, a behavioral testing framework for evaluating NLP models, defines a linguistic capability called ``Negated neutral should still be neutral'' to measure how accurately a \sa model understands that the negated neutral input has neutral sentiment
\cite{marcoACL2020checklist}. 
It requires the \sa model to
output neutral sentiment on the negated neutral inputs. 
Such evaluation methodology avoids the overestimation of the model performance as it measures the model performance on each functionality.
% ,and the separate model performance
% explain distribution of the performance over the linguistic capabilities
In the end, testing through linguistic capabilities provides not only the overall model performance, but also the malfunction facets of the model.

 However, existing linguistic capability-based testing approaches rely on manually generated input templates, which need to be preset before generating tests. 
 As a result, these templates are limited in their sentence
structures. This restricts existing approaches' ability to comprehensively test linguistic capabilities.

@@

To address this issue, we present \tool,
an automated linguistic capability-based testing framework for NLP models.  
The goal of \tool is to generate a diverse linguistic capability-based test suite for a given model, both in terms of covering diverse linguistic structures and diverse model behaviors.
To achieve the goal, \tool needs to address two challenges.
\textit{(i) Capability-based Assessment.} Each test case should be automatically categorized into a \lc;
and \textit{(ii) Automated Test Oracle.} The label of each test case should be automatically
  and accurately defined.
% \end{description}

%\noindent \textbf{C1.} The first challenge comes with the second and third challenges of maintaining the linguistic capability and label of the test sentence while making the syntactic structure of the sentence diverse. To address this challenge, \tool establishes specifications for evaluating a linguistic capability and searches suitable sentences that satisfy the specification from existing public dataset. In this process, \tool generates new inputs by mutating the searched sentences, used as seed inputs. \tool expands the grammar structure of a seed input and determines its available \pos to maintain structural naturalness.

% After that, to hold contextual naturalness of the mutated
% inputs, the fuzzer completes the expanded new structures via
% data-driven context-aware word suggestion. Additionally,
% sentiment-independent words in the inputs are replaced with rule-based
% word suggestion. Further, we address the manual input generation
% process by automating the process mentioned above.

\noindent \textbf{Capability-based Assessment.} The suitability of test sentences for evaluating NLP
models on a \lc is determined by its relevancy to this
\lc. It is challenging to maintain relevancy between a test sentence and its \lc when transforming and expanding the test sentence.
This is because the \lc is defined on a specific 
mixture of syntax and semantics of the sentence. Due to the inherent ambiguity of natural language sentences, there exists no automatic way to check the consistency of each sentence with the semantics specified in the corresponding \lc. To address this difficulty, \tool defines specifications (search rules
and transformation templates) of \lcs, and analyzes the parse
tree of the seed sentence to identify possible expansion.
%% \tool adopts behavioral model testing method by
%% introducing multiple behaviors from \lcs on \sa tasks and generating
%% inputs relevant to each \lc. Each test case provides not only
%% correctness of model prediction on it, but \lc on the test case also
%% gives information about classification on the model
%% performance. Therefore assigning appropriate \lc for each test case
%% alleviates overestimation of model testing results.

\noindent \textbf{Automated Test Oracle.}   For NLP model testing, test oracle is usually determined by understanding the semantics of texts and requires domain knowledge of different NLP tasks.
Thus, the current testing practice requires manual efforts to create the test oracles of the \ho data. The manual work is costly in terms of time consumption. Therefore, it necessitates automated test oracle generation for improving the testing process of NLP models. 
% % validity of the automation is ensured by the
% % correctness of test oracle.
 However, automatically generating the  test oracles remains one of the main challenges in NLP software testing~\cite{huang2022aeon}. 
 A few metamorphic testing approaches have been proposed in image recognition domain, but they require understanding the characteristics of metamorphic relations between inputs and outputs, which require domain expertise and non-trivial manual efforts ~\cite{segura2016metamorphictest}. In addition, it is even more challenging to design metamorphic relations in textual data because the semantics of nature language sentences can be greatly changed even by a slight perturbation to the sentences.
 %Along with this, motivated by software metamorphic testing approach~\cite{segura2016metamorphictest}, we address the challenge by specifying linguistic rules determined by input-output relations, and we only accept validated test oracles by the rules.

% Therefore, generation of test case ought to ensure the
%correctness of its oracle allocation.
In this work, as a first step, we consider \sa as the NLP task for the models under test. 
\tool obtains the
appropriate test oracle by implementing domain-specific
knowledge on the word sentiment dataset and word suggestion model for validating the generated sentence and its oracle.

>>>>

\jl{To address this limitation, increasing number of evaluation
  methodologies has been conducted to measure and evaluate the a NLP
  model in different aspects. First, the NLP model can be fragile to
  adversarial attacks, and it decreases trustworthiness of the NLP
  model. Thus, robustness on adversarial examples are evaluated by
  generating the examples using perturbations to an input
  text.~\cite{morris2020textattack,zang2020sememepso,alzantot2018genadvexp,linyang2020bertattack,udeshi2019ogma}. Second,
  fairness, specific bias in an NLP model toward certain subpopulation
  groups such as gender and races, has been
  measured~\cite{soremekun2020astraea,udeshi2018fairnesstest}. Additionally,
  methodologies to assess an NLP model on a set of \lcs have been
  introduced~\cite{marcoACL2020checklist,rottger2020hatecheck}. Typically,
  it describes certain type of inputs and outputs observed in real
  world for the target NLP task. A \lc includes comprehensive
  functionalities of the input and output behavior for an NLP model,
  and various task-specific model behaviors are evaluated compared
  with the aforementioned evaluation metrics. Therefore,}
% , ranging
% from simple to complicated behaviors.
Testing the linguistic capabilities allows the model developers to
better understand the capabilities and potential issues of the NLP
models.  For example, \Cklst~\cite{marcoACL2020checklist}, a
behavioral testing framework for evaluating NLP models, defines a
linguistic capability called ``Negated neutral should still be
neutral'' to measure how accurately a \sa model understands that the
negated neutral input has neutral sentiment
\cite{marcoACL2020checklist}.  It requires the \sa model to output
neutral sentiment on the negated neutral inputs.  Such evaluation
methodology avoids the overestimation of the model performance as it
measures the model performance on each functionality.
% ,and the separate model performance
% explain distribution of the performance over the linguistic capabilities
In the end, testing through linguistic capabilities provides not only
the overall model performance, but also the malfunction facets of the
model.

However, existing linguistic capability-based testing approaches
generate inputs only relying on word substitution, and need to be
preset before generating tests.  As a result, these templates are
limited in their sentence structures. This restricts existing
approaches' ability to comprehensively test linguistic capabilities.

@@

To address this issue, we present \tool, an automated linguistic
capability-based testing framework for NLP models.  The goal of \tool
is to generate a diverse linguistic capability-based test suite for a
given model, both in terms of covering diverse linguistic structures
and diverse model behaviors.  To achieve the goal, \tool needs to
address two challenges.

\textit{(i) Capability-based text generation.}
Each test case should be automatically categorized into a \lc; and
\textit{(ii) Automated Test Oracle.} The label of each test case
should be automatically and accurately defined.

\noindent \textbf{Capability-based text generation.} The suitability
of test sentences for evaluating NLP models on a \lc is determined by
its relevancy to this \lc. It is challenging to maintain relevancy
between a test sentence and its \lc when transforming and expanding
the test sentence. This is because the \lc is defined on a specific
mixture of syntax and semantics of the sentence. Due to the inherent
ambiguity of natural language sentences, there exists no automatic way
to check the consistency of each sentence with the semantics and
structures specified in the corresponding \lc. \tool defines
specifications (search rules and transformation templates) from \lcs,
and generates seed test cases by searching and transforming test cases
that conform to the specifications. In addition, \tool analyzes the
parse tree of the seed sentence to identify possible expansion and
validates coherence of \lc relevancy between the seed and its
expansion.

\noindent \textbf{Automated Test Oracle.}  For NLP model testing, test
oracle is usually determined by understanding the semantics of texts
and requires domain knowledge of different NLP tasks.  Thus, the
current testing practice requires manual efforts to create the test
oracles of the \ho data. The manual work is costly in terms of time
consumption. Therefore, it necessitates automated test oracle
generation for improving the testing process of NLP models.  However,
automatically generating the test oracles remains one of the main
challenges in NLP software testing~\cite{huang2022aeon}.  A few
metamorphic testing approaches have been proposed in image recognition
domain, but they require understanding the characteristics of
metamorphic relations between inputs and outputs, which require domain
expertise and non-trivial manual efforts
~\cite{segura2016metamorphictest}. In addition, it is even more
challenging to design metamorphic relations in textual data because
the semantics of nature language sentences can be greatly changed even
by a slight perturbation to the sentences.  Motivated by software
metamorphic testing approach~\cite{segura2016metamorphictest}, we
address the challenge by only accepting validated test oracles by the
\lc specification and linguistic rules determined by input-output
relations.

% Therefore, generation of test case ought to ensure the
%correctness of its oracle allocation.
In this work, as a first step, we consider \sa and \hsd as the NLP
tasks for the models under test.  \tool obtains the appropriate test
oracle by implementing domain-specific knowledge on the word sentiment
dataset and word suggestion model for validating the generated
sentence and its oracle.

### 2023-01-13 paper/fse23/related.tex
intro

\MyPara{Text Generation}
%
Different data augmentation approaches have been proposed for text
generation. One of methods in this category is easy data augmentation
(EDA)~\cite{weizou2019eda}. It uses simple operations such as synonym
replacement and random swap. Coulombe~\cite{coulombe2018textaugapi}
implements rule-based transformations using regular expressions for
text transformations such as insertion of spelling mistakes, data
alterations, entity names, and abbreviations. Mixup interpolation,
which was introduced in computer vision category, was implemented in
text domain~\cite{hongyu2019dataaugmixup}. Learning-based language
models has been used to generate text data for multiple
tasks~\cite{qizhe2019unsupdataaug, ateret2019lambda}. While these
approaches focus on obtaining synthetic training data, not testing NLP
model over \lcs,
% they are not able to be used for the capability testing of the NLP model, 
our work generates text data for testing NLP model over \lcs.  In
addition, adversarial text generation methods has been used for
attacking NLP
model~\cite{morris2020textattack,zang2020sememepso,alzantot2018genadvexp,linyang2020bertattack}. Morris \textit{\etal}~\cite{morris2020textattack}
transform input text by leveraging existing methods such as insertion,
deletion and swap of word/character within the similarity-based
constraints. Zang \textit{\etal}~\cite{zang2020sememepso} finds
sememe-based synonyms and generates adversarial examples by the word
substitution. Alzantot \textit{\etal}~\cite{alzantot2018genadvexp}
substitutes word in an input text with its neighboring words in
embedding space that fools an NLP
model. Linyang \textit{\etal}~\cite{linyang2020bertattack} generates
word substitutes with BERT predicted words for attacking an NLP
model. Also, Grammar-based text generation has been also
introduced~\cite{udeshi2019ogma,soremekun2020astraea,maIJCAI2020mtnlp}.
Ogma~\cite{udeshi2019ogma} and Astraea~\cite{soremekun2020astraea}
leverages predefined grammar, and search adversarial examples by
replacing word in an input text with another one that conforms to same
production rule in the grammar until the replacement leads to error in
the model. Ma \textit{\etal}~\cite{maIJCAI2020mtnlp} applies a simple
expansion that inserts an adjective word before noun besides to
replacement of analogy word found in embedding space. Our work
generates text data does not only rely on the word substitution, but
also changes diverse structures from the input text. In addition, we
generates text conforming to same context of input text and its
corresponding \lc by investigating its semantic and syntax.

>>>>

\MyPara{Text Generation}
%
Different data augmentation approaches have been proposed for text
generation. One of methods in this category is easy data augmentation
(EDA)~\cite{weizou2019eda}. It uses simple operations such as synonym
replacement and random swap. Coulombe~\cite{coulombe2018textaugapi}
implements rule-based transformations using regular expressions for
text transformations such as insertion of spelling mistakes, data
alterations, entity names, and abbreviations. Mixup interpolation,
which was introduced in computer vision category, was implemented in
text domain~\cite{hongyu2019dataaugmixup}. Learning-based language
models has been used to generate text data for multiple
tasks~\cite{qizhe2019unsupdataaug, ateret2019lambda}. While these
approaches focus on obtaining synthetic training data, not testing NLP
model over \lcs,
% they are not able to be used for the capability testing of the NLP model, 
our work generates text data for testing NLP model over \lcs.  In
addition, text generation methods has been used for attacking NLP
model~\cite{morris2020textattack,zang2020sememepso,alzantot2018genadvexp,linyang2020bertattack}.
Morris \textit{\etal}~\cite{morris2020textattack} transform input text
by leveraging existing methods such as insertion, deletion and swap of
word/character within the similarity-based
constraints. Zang \textit{\etal}~\cite{zang2020sememepso} finds
sememe-based synonyms and generates adversarial examples by the word
substitution. Alzantot \textit{\etal}~\cite{alzantot2018genadvexp}
substitutes word in an input text with its neighboring words in
embedding space that fools an NLP
model. Linyang \textit{\etal}~\cite{linyang2020bertattack} generates
word substitutes with BERT predicted words for attacking an NLP
model. Also, Grammar-based text generation has been also
introduced~\cite{udeshi2019ogma,soremekun2020astraea,maIJCAI2020mtnlp}.
Ogma~\cite{udeshi2019ogma} and Astraea~\cite{soremekun2020astraea}
leverages predefined grammar, and search adversarial examples by
replacing word in an input text with another one that conforms to same
production rule in the grammar until the replacement leads to error in
the model. Ma \textit{\etal}~\cite{maIJCAI2020mtnlp} applies a simple
expansion that inserts an adjective word before noun besides to
replacement of analogy word found in embedding space. Our work
generates text data does not only rely on the word substitution, but
also changes diverse structures from the input text. In addition, we
investigate input semantic and syntax to generate text conforming to
the same context as input text and its corresponding \lc.
